---
title: "1_extract_lng_data"
author: "Lia Domke"
date: "1/5/2026"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# read in libraries
```{r}
# detach("package:EMAdownload", unload = T)
# remotes::install_github("afsc-ema/EMAdownload", quiet = F, force = T, dependencies=TRUE, build_vignettes=TRUE)
# rm(list = c("evnt", "fsh", "event_parameters", "taxa"))


library(EMAdownload)
library(tidyverse)
library(ggplot2)
library(ggridges)
library(ggmap)
library(maps)
library(mapdata)
library(mapproj)
# mapping
library(marmap)
# library(PBSmapping)
```

custom graph theme
```{r}
plot_theme <- function() {
  theme_bw(base_size = 16, base_family = "Avenir") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="white", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            strip.background = element_rect(colour = "NA", fill = "grey96"))
}
```


# Data pull from AKFIN
```{r}
# pull the tax data so we can get the species tsn for Pacific Cod
# tax <- EMAdownload::get_ema_taxonomy() %>%
#   filter(scientific_name == "Gadus macrocephalus")

# Question for Mariela - do we care about if the data come from a surface trawl or are oblique trawls acceptable? 
# for now we'll include only surface trawls done with can trawls for now
# we use join_event_fish because it pulls out all SPECIMEN level data (so no catch level data). This also takes some time to run
# because if we pull all years its a lot of information. 

# we can do catch weighting, where we 
pcod <- join_event_fish(start_year = 2003, end_year = 2026, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN")

# pcod <- read.csv("data/pcod_lengths_akfin_022726.csv") %>%
#   dplyr::select(-c(X, X.1))

# write.csv(pcod, "data/pcod_lengths_akfin_022726.csv", row.names = F)

pcod_catch <- join_event_catch(start_year = 2003, end_year = 2025, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN", catch0 = TRUE)

#write.csv(pcod_catch, "data/pcod_catch_akfin_011526.csv")
#pcod_catch <- read.csv("data/pcod_catch_akfin_011526.csv")

table(pcod_catch$lhs_code)
```

Quick look at the data  
```{r}
head(pcod)
table(pcod$sample_year)

table(pcod$length_type) # two length type SL/TL
table(pcod$lhs_code) # multiple life histories here, instead of using the lhs code we'll use a reasonable length cutoff
table(pcod$lhs_code, pcod$length_type) # some tl are a1+ 
# we can only convert TL to SL for pcod under 100 mm

table(pcod$length < 100, pcod$lhs_code) # there are some a0 greater than 100
table(pcod$length < 100, pcod$length_type) # there are very few pcods that are

table(pcod$length_type, pcod$sample_year)

# Question, what is a reasonable length cutoff?
```

Quick view of length dsitributions AS IS
```{r}
pcod %>%
  filter(length < 200) %>%
  ggplot() +
  geom_density_ridges(aes(x = length, y = as.factor(sample_year), group = sample_year, fill = as.factor(sample_year))) +
  facet_grid(cols = vars(region))

pcod %>%
  ggplot() +
  geom_histogram(aes(x = length, fill = lhs_code))

test <- filter(pcod, length < 150 & lhs_code %in% c("A0", "A1+")) %>%
  filter(sample_year %in% c(2006, 2007, 2009, 2011))

#write.csv(test, "confusing_lhs_pcod_lng.csv")

ggplot(data = test) +
  geom_histogram(aes(x = length, fill = lhs_code)) +
  facet_wrap(~sample_year)

unique(test$sample_year)

conc.size <-pcod_catch %>%
  filter(sample_year %in% c(2006, 2007, 2009, 2011),
         total_catch_number > 0)

table(conc.size$lhs_code)
```

What is the annual sample counts by station in NBS & SEBS
```{r}
# because there are master station names that are inconsistent etc over years that haven't been fixed here we recalculate master station name
pcod$master_station_name <- paste(plyr::round_any(pcod$eq_latitude,0.25),plyr::round_any(pcod$eq_longitude,0.25),sep="")

table(is.na(pcod$master_station_name))# great all gone 

# use master station name which is a rounded version of lat/lon together
pcod %>%
  filter(length < 200) %>%
  filter(region == "NBS") %>%
  group_by(region, sample_year, master_station_name) %>%
  dplyr::summarize(num = n()) %>%
  ggplot() +
  geom_col(aes(x = master_station_name, y = num)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(rows = vars(sample_year))

pcod %>%
  filter(length < 200) %>%
  filter(region == "SEBS") %>%
  group_by(region, sample_year, master_station_name) %>%
  dplyr::summarize(num = n()) %>%
  ggplot() +
  geom_col(aes(x = master_station_name, y = num)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(rows = vars(sample_year))

table(pcod$sample_year, pcod$region)
table(pcod$sample_year, pcod$master_station_name)
```


create base map
```{r}
US <- map_data("world2Hires",region="USA")
AK <- subset(US,subregion == "Alaska")
USSR <- map_data("world2Hires",region="USSR")
NBS <- rbind(AK,USSR)
#transforms longitudes (you could also transform the longs in the data file)
NBS$long <- NBS$long-360
#get bathymetry data [note it's a good idea to save this as a .rds file.  This command queries a database and it can be down sometimes]
bathline <- getNOAA.bathy(-150,-180,50,65)
```


```{r}

ggplot() +
    geom_polygon(data=NBS,aes(x=long,y=lat,group=group),fill="wheat3") +
    geom_contour (data=bathline,aes(x=x,y=y,z=z),breaks=c(-50),size=c(0.3),colour="seagreen",alpha=0.2) +
    geom_contour (data=bathline,aes(x=x,y=y,z=z),breaks=c(-200),size=c(0.3),colour="green3",alpha=0.2) +
    annotate( geom = "text", x = -161.8, y = 57.4, label = "50m", 
              color = "black", size = 2 )+
    annotate( geom = "text", x = -168.1, y = 56, label = "200m", 
              color = "black", size = 2 )+
    scale_y_continuous(breaks=c(56,58,60, 62, 64, 66)) +
    scale_x_continuous(breaks=c(-175, -170,-165,-160, -155)) +
    coord_sf(crs="EPSG:3571",default_crs=sf::st_crs(4326),xlim=c(-175,-155),ylim=c(55,66)) +
    geom_point(data = {filter(pcod, length < 200)},
             aes(x = eq_longitude, y = eq_latitude)) +
  facet_wrap(~sample_year)
```

Note we have some years where Pacific cod were measurred both in TL or SL. 
The convention in recent years is to use SL. To convert historic information for individuals between 30-100 mm:
The conversions for Pacific cod (Gadus macrocephalus) in the 30-100mm range are

TL(mm) = 0.26250 + 1.0769*SL(mm)

SL(mm) = 0.06258 + 0.92843*TL(mm) 

# convert TL to SL
Here we want to convert the TL to SL for all fish < 100 mm
```{r}
pcod_sl_sub <- pcod %>%
  filter(length < 150 & length_type == "TL") %>%
  mutate(length = 0.06258 + 0.92843*length,
         length_type = "SL")

table(pcod$length < 150, pcod$length_type)
table(pcod_sl_sub$length_type)

pcod %>%
  filter(length < 150 & length_type == "TL") %>%
  mutate(length_adj = 0.06258 + 0.92843*length) %>%
  ggplot() +
  geom_density_ridges(aes(x = length, y = as.factor(sample_year), group = sample_year), fill = "green", alpha = 0.8) +
  geom_density_ridges(aes(x = length_adj, y = as.factor(sample_year), group = sample_year), fill = "blue", alpha = 0.8)

pcod_sl <- pcod %>%
  filter(!(length < 150 & length_type == "TL"))

nrow(pcod_sl); nrow(pcod_sl_sub); nrow(pcod)
nrow(pcod_sl) + nrow(pcod_sl_sub)

pcod_sl_full <- rbind(pcod_sl_sub, pcod_sl) 

pcod_sl_full %>%
  filter(length < 150) %>%
  ggplot() +
  geom_density_ridges(aes(x = length, y = as.factor(sample_year), group = sample_year, fill = as.factor(sample_year))) +
  facet_grid(cols = vars(region))
```


Density graph of size frequencies w/ mean length in a dashed line - grouped across years
```{r}
pcod_means <- pcod_sl_full %>%
  filter(length < 150 & length_type == "SL") %>% # this takes the full pcod dataset (nrow 12453) and reduces it to 12346
  group_by(sample_year, region) %>%
  dplyr::summarize(mean_lng = mean(length), 
                   median_lng = median(length),
                   num_lng = n()) %>%
  ungroup() %>%
  complete(sample_year = 2003:2025, region = c("NBS", "SEBS"))

pcod_sl_full %>%
  filter(length < 150 & length_type == "SL",
         sample_year > 2005) %>%
  ggplot() +
  geom_density(aes(x = length, group = as.factor(sample_year), fill = as.factor(sample_year)), alpha = 0.5) +
  geom_vline(data = pcod_means, aes(xintercept = mean_lng, color = as.factor(sample_year)), linetype = "dashed", linewidth = 1.5) +
  facet_wrap(~region) +
  labs(x = "Length (mm)", y = "Density", fill = "Year") +
  plot_theme() +
  guides(color = "none")

  
  
```


Density graph of size frequencies w/ mean length in a dashed line - separated by year
```{r}

pcod_sl_full %>%
  filter(length < 150 & length_type == "SL",
         sample_year > 2005) %>%
  ggplot() +
  geom_density_ridges(aes(x = length, y = as.factor(sample_year), group = as.factor(sample_year), fill = as.factor(sample_year))) +
  geom_vline(data = pcod_means, aes(xintercept = mean_lng, color = as.factor(sample_year)), linetype = "dashed", linewidth = 1.5) +
  facet_wrap(~region) +
  labs(x = "Length (mm)", y = "Density", fill = "Year") +
  theme_bw() +
  guides(color = "none")
  
# What I dont lvoe about this graph is that it doesn't do a good job of conveying actual numbers. 
# For example in NBS 2015 only 8 pcod were lengthed so the mean looks far right of the peak in the density
```

Quick investigation into oceanographic domain

# Fig oceanographic domain with pcod SL density across years
## (MB use this)
```{r}
pcod_means_oceano <- pcod_sl_full %>%
    filter(length < 150 & length_type == "SL",
         sample_year > 2005) %>%
  group_by(oceanographic_domain, sample_year) %>%
  dplyr::summarize(mean_lng = mean(length), 
                   median_lng = median(length),
                   num_lng = n(),
                   sd_lng = sd(length),
                   se_lng = sd_lng/sqrt(num_lng)) %>%
  ungroup() %>%
  complete(oceanographic_domain = c(1,2,3,4),
           sample_year = 2006:2025) %>%
  filter(oceanographic_domain %in% c(1,2)) %>%
  mutate(oceanographic_label = ifelse(oceanographic_domain == 1, "Inner shelf",
                                      ifelse(oceanographic_domain == 2, "Middle shelf", NA)))


amss_Fig1 <- 
pcod_sl_full %>%
  filter(length < 150 & length_type == "SL",
         sample_year > 2005, oceanographic_domain %in% c(1,2),
         sample_year < 2025,
         !(sample_year %in% c(2019))) %>%
  complete(sample_year = 2006:2024) %>% # remove this line if you only want to include years that we decided to include data from.
  # i.e. 2006-2012, 2014-2018, 2021-2024
  mutate(oceanographic_label = ifelse(oceanographic_domain == 1, "Inner shelf",
                                      ifelse(oceanographic_domain == 2, "Middle shelf", NA))) %>%
  ggplot() +
  geom_density_ridges(alpha = 0.6,
                      aes(x = length, y = as.factor(sample_year),
                          fill = as.factor(oceanographic_label))) +
  scale_fill_manual(values = c("darkblue","orange")) +
  labs(x = "Length (mm)", y = "Year", fill = "Oceanographic Domain") +
  theme_bw() +
  theme(axis.line = element_line(color = "black"),
        plot.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(color = "black", fill = NA),
        axis.text = element_text(color = "black"),
        legend.title = element_blank(),
        legend.position = "none", # uncomment to remove legend for saving figure
        text = element_text(size = 32)) # change to 32 or 34 to save figure

amss_Fig1

# ggsave("amssFig1_LengthDensityPlots_axisTextFix_011526.png",plot = amss_Fig1,
#       height = 14,
#       width = 10,
#       dpi = 300,
#       units = "in",
#       path = "figures/")

  
```


Correct Length for day of year and latitude...how does this affect distributions? Is there a better way to apply this correction?

```{r, echo = FALSE}

# # Define YearDay for length data
pcod_sl_full$YearDay <- 
  lubridate::yday(pcod_sl_full$haul_date) # Need sample date data before we can run this...

mod <- glm(log(length) ~ eq_latitude + YearDay, data = pcod_sl_full)
pcod_sl_full$glmLengthLatYday_resid <- resid(mod)
#pcod_sl_full$glmLengthLatYday_predict 
# test_predict <- predict(mod, pcod_sl_full$length, type = "response")
# fitstatsLengthLatYday <- summary(mod)
# fitstatsLengthLatYday

```

What if we used counts rather than density
```{r}
pcod_sl_full %>%
  filter(length < 150 & length_type == "SL",
         sample_year > 2005, oceanographic_domain %in% c(1,2)) %>%
  mutate(oceanographic_label = ifelse(oceanographic_domain == 1, "Inner shelf",
                                      ifelse(oceanographic_domain == 2, "Middle shelf", NA))) %>%
  ggplot() +
  geom_freqpoly(aes(x = length, color = as.factor(oceanographic_label)), linewidth = 1.5) +
  geom_vline(data = pcod_means_oceano, aes(xintercept = mean_lng, color = as.factor(oceanographic_label)), linetype = "dashed", linewidth = 1.5) +
  #geom_ribbon(data = pcod_means_oceano, aes(xmin = mean_lng - se_lng, xmax = mean_lng + se_lng, y = num_lng)) +
  facet_wrap(~sample_year) +
  labs(x = "Length (mm)", y = "Count", color = "Oceanographic Domain") +
  theme_bw()
```


Pcod catch
```{r}
head(pcod_catch)

pcod_catch %>%
  mutate(cpue_num = total_catch_number/effort) %>%
  ggplot() +
  geom_boxplot(aes(y= cpue_num, x = as.factor(sample_year), group = as.factor(sample_year))) +
  facet_wrap(~region, scales = "free")

pcod_catch %>%
  mutate(cpue_num = total_catch_number/effort) %>%
  ggplot() +
  geom_point(aes(y= cpue_num, x = as.factor(sample_year), group = as.factor(sample_year))) +
  facet_wrap(~region, scales = "free")

```

experimental modelling 
```{r}
names(pcod_catch)

pcod_cpue <- pcod_catch %>%
  group_by(sample_year, station_id, lhs_code, gear) %>%
  mutate(cpue = total_catch_number/effort)

lm<- lm(cpue ~ sample_year, data = pcod_cpue)

library(DHARMa)
plot(lm)

residuals <- simulateResiduals(lm)
plotResiduals(residuals)
plotQQunif(residuals)
```

Okay here we will be preparing the data for use 
We want to take the catch and lengthed data and correct the lengths

We were going to set up these length data to catch weight the lengths by the number of indivduals
that were sampled, but i've decided based on trying to do that with the age-0 pollock data
that it creates too large of a dataframe that perhaps doesn't make sense. 
its trying to aggregate both the abundance information and length information into a single analysis when perhaps we have separate question for each one of those... 


Instead let us correct the length information for Pcod

# Correct remaining Pcod

```{r}
nrow(pcod)
head(pcod) # uses surface catch only in nbs and sebs 2003:today
nrow(pcod_catch) # same here
```

```{r}
unique(pcod_catch$tow_type); unique(pcod_catch$gear)

# look for catch 0 positive weight discrepancy & vice versa
discrep <- filter(pcod_catch, total_catch_number == 0 & total_catch_weight != 0)
discrep2 <- filter(pcod_catch, total_catch_number !=0 & total_catch_weight == 0)

# there's no length data for the two stations that were sampled in 2003 (n = 9 and 5; 20030101080 & 20030101007
# checking the paper records there were only catch numbers recorded for those stations and no lengths
# for 20030101080 9 individuals were caught and processed to say they were between 11.2-12 cm (112 - 120 mm)
# for 20030101007 5 were caught and all weren't processed
test<- subset(pcod, station_id %in% discrep2$station_id)

# no lengths taken where the catch was weighed and no catch number stimated, which is what i'd expect. 
test2 <- subset(pcod, station_id %in% discrep$station_id)

# so from these I won't be able to use these stations to estimate information
```



```{r}
## we have two problems that need to be dealt with here
#1. that there are tl measured pcods that should be SL
table(pcod$length_type) # we know we have TL that should be SL
pcod.tl <- filter(pcod, length_type == "TL")
table(pcod.tl$sample_year, pcod.tl$lhs_code)
# how many pcod greater than the 100 mm equation cut off that we think need to be changed
filter(pcod, length_type == "TL" & length > 100 & length < 140) %>%
  nrow(.) # approx 924 instances that will need to be adjusted and perhaps won't be correct, but we don't have a better 
# equation to base it off of. However actual number is likely different because the 140 mm cut off is based on SL
total_lng_cutoff <- 0.26250 + 1.0769*140 # this is the total length cutoff, 151.028

pcod_corrected <- pcod %>%
  mutate(corrected_lng = ifelse(length_type == "TL" & length < total_lng_cutoff,
                         0.06258 + 0.92843*length,
                         ifelse(length_type == "TL" & length > total_lng_cutoff, NA,
                                ifelse(length_type == "SL" & length < 140, length, NA)))) 

pcod_corrected

# 2. there are pcod classified as A1+ that are within the 140 mm range
# quick test to see if there are any pcod less than 140 that migth need to be 
# reclassified
test_a0_pcod_lng <- filter(pcod, length < 140 & lhs_code == "A1+")
table(test_a0_pcod_lng$lhs_code) # all pcod that were less than 140 w/ lhs A1+ were fixed
filter(pcod, length > 140 & lhs_code == "A0") # at least theres no big pcod classified as A0
```

# Output data
```{r}
#write.csv(pcod_corrected, "data/pcod_length_corrected_050426.csv", row.names = F)
```

