---
title: "3_catch_sdmTMB_exploration"
output: html_document
date: "2026-03-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#libraries
```{r}
library(EMAdownload)
library(tidyverse)
library(ggplot2)
library(sdmTMB)
library(patchwork)
```

# data
```{r}
this.year <- format(Sys.Date(), "%Y") %>% as.numeric()

pcod_catch <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN", catch0 = TRUE) %>%
  filter()
# note akfin upload date, if in Jan hasn't been updated to include the correct pcod cut off point for 140 mm
```

go through some of the basic steps for sdmTMB 

- add_utm_columns
- make mesh, make_mes()
- fit model 
    - modelling catch data with effort data
    - for now i think we're more interestd in catch data rather than weightcpue
    - an easy start for distribution would be to use the tweedie model (alt. could use a poisson with a effort offset)
    
```{r}
# check to make sure the format makes sense
head(pcod_catch)
unique(pcod_catch$common_name)
range(pcod_catch$total_catch_number)

table(pcod_catch$total_catch_number > 0) # lots of zeros 
table(pcod_catch$sample_year, pcod_catch$total_catch_number >0)

hist(pcod_catch$total_catch_number) # kinda looks zero inflated :(
# for now lets move forward with it the way it is and see how model fits look

table(pcod_catch$lhs_code, pcod_catch$total_catch_number > 0)

pcod_catch_a0 <- filter(pcod_catch, lhs_code == "A0")
```

## add utms & mesh
```{r}
catch.32603 <- add_utm_columns(pcod_catch_a0,
                ll_names = c("eq_longitude", "eq_latitude")) %>%
  dplyr::select(c(sample_year, station_id, master_station_name:region, eq_time:eq_longitude, effort, effort_units, 
                  species_tsn:total_catch_weight, X, Y)) %>%
  mutate(cpue = total_catch_number / effort)
# this uses CRS 32603

wgs <- ggplot() +
  geom_point(aes(eq_longitude, eq_latitude), data = catch.32603)

utm <- ggplot() +
  geom_point(aes(X, Y), data = catch.32603)

wgs + utm

# this is somewhat of a high number of knots (I think) but lets try it out for now
mesh <- make_mesh(catch.32603, c("X", "Y"), n_knots = 80)
plot(mesh)
# might want to consider some kind of barrier here to make sure that we're not trying to make a correlation matrix thingy with places with no data that don't make sense
```

## fit model

this is a first pass to see if this converges but these are why i think i'm making these decisions
```{r}
# this model estaimtes: independent year effects (because of sample_year in model formula),
# a
fit1 <- sdmTMB(
  data = catch.32603,
  # we're curious about the total_catch_number (abundance) that may vary by time (so estimating a diff parameter for each year) 
  total_catch_number ~ 0 + as.factor(sample_year),
  # this is the mesh we created above that (we think) is a finer scale mesh, can be coarser if conversion issues arise
  mesh = mesh,
  # since we have spatio temporal fields on (w/ iid) we need to indicate what our time column is
  time = "sample_year",
  # tweedie family is a good starting point generally used with continuous and/or zero inflated biomass data the
  # catch data is perhaps not the most appropriate but we'll start here, maybe poisson with a catch offset would be better?
  family = tweedie(link = "log"), 
  # spatiotmeporal field is iid meaning each year is independently estimated 
  # this is another spot we could loosen up on flexibility if it doesn't converge (i.e., ar1 or rw)
  spatiotemporal = "IID",
  # anisotropy v isotrophy this is the decorrelation assumptions with Matern distance. 
  # basically it assume that the decorrelation dist the same E-W and N-S however in the BS 
  # we would expect that there would be greater correlation along shelf (NW/SE) than across shelf E-W
  # in general keep unless its unestimable 
  anisotropy = TRUE
)

fit1

# things of interest here, 
#' dispersion - how much variability after account for main effects and spatial/spatiotemporal 
#' matern range - indicates if smooth or patchy
#' spatiial SD if high means theres strong peaks / valleys in spatial fields, if 0 no spatial variability
#' 
tidy(fit1, "ran_pars")
#' range smooth or patchy (but which is which)
```

## model checks
```{r}
# looking for approximately normal along the line
catch.32603$resid <- residuals(fit1, type = "mle-mvn")
qqnorm(catch.32603$resid)
abline(0,1)

hist(catch.32603$resid)

# spatial residuals
# here we're revuing the data to make sure we're not seeing patterns of clustering of
# positive and negative residuals together
ggplot() +
  geom_point(data = catch.32603, aes(X, Y, color = resid)) +
  scale_color_gradient2() +
  facet_wrap(~sample_year)
# kinda looks okay?

# passes the standard sanity checks
sanity(fit1)
```

## predict obs
```{r}
preds <- predict(fit1)

#' est: fixed + spatial + spatiotemporal (use this for maps and indices)
#' est_non_rf - fixed effects only, use this to understand the covariate-driven signal (in our case here this would
#' be the temporal effects)
#' omega_s - spatial random effects use this to see persistent spatial deviations 
#' epsilon_st - spatiotemporal random effects, will tell you/show year-specific anomalies


# i think this is the spatiotemporal random effects fields
preds %>% 
  ggplot(aes(x=X, y=Y, color=epsilon_st)) + 
  #theme_linedraw() +
  scale_color_viridis_c() +
  geom_point() +
  facet_wrap(~sample_year) +
  ggtitle("Spatiotemporal RE (link space)",subtitle="Spatiotemporal Model")

# to includes overall estimate that includes the main effects

preds %>%
  ggplot(aes(x =X, y = Y, color = est)) +
  scale_color_viridis_c() +
  geom_point() +
  facet_wrap(~sample_year)
  
```

## prediction grids
```{r}
library(sf)
library(sfheaders)
library(concaveman)
# steps : 
# create sf object, converts to metric
catch.sf <- catch.32603 %>%
  dplyr::select(c(X, Y, eq_latitude, eq_longitude)) %>%
  mutate(X = X * 1000, Y = Y * 1000) %>% # convert to meters for crs 32606 (but we're using 32603 i think thats ok)
  st_as_sf(coords = c("X", "Y"), crs = 32603)

# check to make sure it makes sense
head(catch.sf$geometry)
# create convex hull 
convex_hull <- concaveman(catch.sf, concavity = 3)
plot(convex_hull)

# add buffer
convex_buff <- st_buffer(convex_hull, dist = 15e3) # dist in meters (i think) so this is 15 km buffer
plot(convex_buff)

# make grid, created a cell size of 10,000 x 10,000 m (10 x 10 km) with poitns in center
# need to check units on this... 
grid <- st_make_grid(convex_buff, cellsize = 10e3, what = "centers") %>%
  st_intersection(convex_buff)
plot(grid)

# change back to df, this wasn't straightforward for me...
bs.grid <- do.call(rbind, st_geometry(grid)) %>%
  as.tibble() %>% setNames(c("X", "Y")) %>%
  # convert df m back to km
  mutate(X = X/1e3, Y = Y/1e3)

plot(bs.grid)

# replicate the df grid for each year (or other variables we include either spatio temporal or temporal) depth, temp etc
bs.grid.yrs <- replicate_df(bs.grid, "sample_year", unique(catch.32603$sample_year)) # does this across 2003:2025, but no 2020

```

## predict across grid
```{r}
preds <- predict(fit1, newdata = bs.grid.yrs, return_tmb_object = F)

preds %>%
  ggplot(aes(X, Y, fill =est)) +
  scale_fill_viridis_c() +
  geom_raster() +
  facet_wrap(~sample_year)
```


## get indicies
jk we're going to wait on this cause we want to make a prediction grid first
```{r}
#index_preds <- predict(fit1, return_tmb_object = TRUE)
?get_cog
#get_index(index_preds)

index_preds <- predict(fit1, newdata = bs.grid.yrs, return_tmb_object = TRUE)
index <- get_index(index_preds)

ggplot(index) +
  geom_point(aes(sample_year, est)) +
  geom_errorbar(aes(sample_year, ymin = lwr, ymax = upr))

cog <- get_cog(index_preds)
head(cog)

```


