---
title: "3_catch_sdmTMB_exploration"
output: html_document
date: "2026-03-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#libraries
```{r}
library(EMAdownload)
library(tidyverse)
library(ggplot2)
library(sdmTMB)
library(patchwork)
library(sf)
library(marmap)
library(DHARMa)
library(ggplotify)
library(here)
library(lubridate)
```

# data
```{r}
this.year <- year(Sys.Date())

pcod_catch <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN", catch0 = TRUE) 
# note akfin upload date, if in the upload date says some time in Jan 2026 then the pcod data hasn't been updated to include the correct pcod cut off point for 140 mm

# pcod_catch_nets <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
#                 trawl_method = "O", gear = "NETS156", catch0 = TRUE) 
```

# create base/bathy maps
```{r}
# make sure the base map / bathy is created with the advanced sf library 
sf_use_s2(TRUE)

# need to use world not world2hires cause that one is hella old and doesn't plot correctly
bs_map <- maps::map("world", regions = c("USA:Alaska", "Russia"), 
                    fill = TRUE, plot = FALSE, wrap = c(0, 360)) %>%
  st_as_sf(crs = 4326) 

# Use bounding box, have to add 360 to the bbox coords
# xmin = 185 (-175), xmax = 202 (-155)
bbox_360 <- st_bbox(c(xmin = 180, xmax = 210, ymin = 50, ymax = 68), 
                    crs = st_crs(bs_map))

# crop 
# Using st_make_valid helps w/ messy polys
bs_base <- bs_map %>% 
  st_make_valid() %>% 
  st_intersection(st_as_sfc(bbox_360)) %>%
  st_shift_longitude()

bs_croppedUTM <- bs_base %>%
  st_transform(crs = 32602) # Move to UTM Zone 2N

#read in bathymetry file.  
bathline <- getNOAA.bathy(-180,-155,53,68)

bathline.xyz <- as.xyz(bathline)
```


go through some of the basic steps for sdmTMB 

- add_utm_columns
- make mesh, make_mes()
- fit model 
    - modelling catch data with effort data
    - for now i think we're more interestd in catch data rather than weightcpue
    - an easy start for distribution would be to use the tweedie model (alt. could use a poisson with a effort offset)
    
```{r}
# check to make sure the format makes sense
head(pcod_catch)
unique(pcod_catch$common_name)
range(pcod_catch$total_catch_number)

table(pcod_catch$total_catch_number > 0) # lots of zeros 
table(pcod_catch$sample_year, pcod_catch$total_catch_number >0) 
# proportionally very few stations where pcod were caught. 

hist(pcod_catch$total_catch_number) # kinda looks zero inflated :(
# for now lets move forward with it the way it is and see how model fits look

table(pcod_catch$lhs_code, pcod_catch$total_catch_number > 0)

pcod_catch_a0 <- filter(pcod_catch, lhs_code == "A0")
table(pcod_catch_a0$sample_year, pcod_catch_a0$total_catch_number >0) 

filter(pcod_catch_a0, is.na(effort)) # mising effort at DY1408, station 20140308022

sample.size <-pcod_catch_a0 %>%
  group_by(sample_year, region) %>%
  dplyr::summarise(total_catch = sum(total_catch_number),
                   total_weight = sum(total_catch_weight))

pcod_catch_a0 %>%
  group_by(region) %>%
  dplyr::summarise(mean_catch = mean(total_catch_number))
```

## clean data 
```{r}
pcod_catch_a0 <- pcod_catch %>%
  filter(lhs_code == "A0",
         total_catch_weight != -999,
         !(is.na(effort))) %>%
  mutate(log_effort = log(effort),
         doy = yday(haul_date),
         median_doy = median(doy),
         scale_doy = doy - median_doy,
         mean_doy = mean(doy),
         sd_doy = sd(doy),
         scale_doy2 = (doy - mean_doy)/sd_doy,
         scaled_median = (median_doy - mean_doy) / sd_doy)
```


## add utm columns
```{r}
catch.32603 <- pcod_catch_a0 %>%
  add_utm_columns(ll_names = c("eq_longitude", "eq_latitude"),
                ll_crs = 4326,
                units = "km",
                utm_names = c("X", "Y")) %>%
  dplyr::select(c(sample_year, station_id, master_station_name:region, 
                  eq_time:eq_longitude, effort, effort_units, 
                  species_tsn:total_catch_weight, doy, scale_doy2, scaled_median, X, Y)) %>%
  mutate(cpue = total_catch_number / effort,
         total_catch_weight = total_catch_weight / 1000)
# this uses CRS 32603
```

## make mesh
```{r}
# this is somewhat of a high number of knots (I think) but lets try it out for now
mesh <- make_mesh(catch.32603, 
                  c("X", "Y"), 
                  n_knots = 200)
plot(mesh)
# might want to consider some kind of barrier here to make sure that we're not trying to make a correlation matrix thingy with places with no data that don't make sense
```

## fit model

this is a first pass to see if this converges but these are why i think i'm making these decisions

? How do we make decisions about the spatial versus spatiotemporal fields? 
first we have random effects for a given year and location that are the summation of the spatial and spatiotemporal random effects. 
- the spatial field is a static spatial pattern consistent across years (+/- the sd (or sigma_0))
- spatiotemporal field provides the year specific deviations, which when it is iid are idependently estimated (+/- sd or sigma_E). 

potential combinations for spatial fields
1. spatial on + spatiotemporal iid: this assumes some baseline of species distribution (spatial), but that each year for some temporally varying factor that causes varying in the species distirbution. From a model like this we can look at the constant spatial pattern which is from pooling all years of data (ws). Then the annual field (e subscript st) represent the deviations (or spatial anomolies) of each year

2. spatial off + spatiotemporal iid: because spatial is off there's no baseline home ranges of the species. Each year the model estimates where the species distribution is de-novo from that years data. There's no interannual variation (i.e., 2009 isn't related to spatial distrb with 2010). 

Should you keep the spatial "ON"? 
- if the species have static habitat preferences then the average map is pretty consistent year after year. 
- if there are data poor years or missing station. Then you can borrow strength from data rich years and those missing data areas (NOTE missing data NOT zero occurence) are filled in with the "average" field. 

When do you remove spatial field, ie spatial "OFF"?
- When AIC prefers the model without spatial field
- sigma_O decomposses to 0 meaning there's no variation in the spatial field and it should be removed. 
- if species are nomadic/no driven by habitat prefereences 
```{r}
# this model estimates: independent year effects 
# (because of sample_year in model formula),
fit1 <- sdmTMB(
  data = catch.32603,
  # we're curious about the total_catch_number (abundance) that may vary by time (so estimating a diff parameter for each year) 
  total_catch_number ~ 1 + as.factor(sample_year) + scale_doy2,
  offset = log(catch.32603$effort),
  # this is the mesh we created above that (we think) is a finer scale mesh, can be coarser if conversion issues arise
  mesh = mesh,
  # since we have spatio temporal fields on (w/ iid) we need to indicate what our time column is
  time = "sample_year",
  # tweedie family is a good starting point generally used with continuous and/or zero inflated biomass data the
  # catch data is perhaps not the most appropriate but we'll start here, maybe poisson with a catch offset would be better?
  family = tweedie(link = "log"), 
  # spatiotmeporal field is iid meaning each year is independently estimated 
  # this is another spot we could loosen up on flexibility if it doesn't converge (i.e., ar1 or rw)
  spatiotemporal = "IID",
  # spatial is on when spatiotemporal is off, but it can be on with spatiotemporal
  spatial = "on", 
  # anisotropy v isotrophy this is the decorrelation assumptions with Matern distance. 
  # basically it assume that the decorrelation dist the same E-W and N-S however in the BS 
  # we would expect that there would be greater correlation along shelf (NW/SE) than across shelf E-W
  # in general keep unless its unestimable 
  anisotropy = TRUE
)

```

Lets quickly compare with other models with spatial off for example
```{r}
fit2 <- sdmTMB(
  data = catch.32603,
  total_catch_number ~ 1 + as.factor(sample_year) + scale_doy2,
  offset = log(catch.32603$effort),
  mesh = mesh,
  time = "sample_year",
  family = tweedie(link = "log"), 
  spatiotemporal = "IID",
  spatial = "off", 
  anisotropy = TRUE
)

# what if we changed the distributional form

fit3 <- sdmTMB(
  data = catch.32603,
  total_catch_number ~ 1 + as.factor(sample_year) + scale_doy2,
  offset = log(catch.32603$effort),
  mesh = mesh,
  time = "sample_year",
  family = delta_gamma(link1 = "log", link2 = "log"), 
  spatiotemporal = list("iid", "iid"),
  spatial = list("on", "on"), 
  anisotropy = TRUE
) # okay noted... doesn't even converge

# slightly better fit with spatial field on. Best way of evaluating this would be cross 
AIC(fit1); AIC(fit2) # validation
```

### cross validation

This takes awhile to run you can skip it if you're not looking at cv
```{r eval=FALSE, include=FALSE}
fold_vector <- as.numeric(as.factor(catch.32603$sample_year))
catch.32603$log_effect <- log(catch.32603$effort)
# open the parallel processing 
TMB::openmp(n = 1)
# set up paralell processing
future::plan(future::multisession, workers = 4)
# run models. 

fit2.cv <- sdmTMB_cv(
  data = catch.32603,
  total_catch_number ~ 1 + scale_doy2,
  offset = "log_effort",
  mesh = mesh,
  time = "sample_year",
  family = tweedie(link = "log"), 
  spatiotemporal = "IID",
  spatial = "off", 
  anisotropy = TRUE, 
  fold_ids = fold_vector
) # no warnings thats good, can't run sanity cause its like a list of models

fit1.cv <- sdmTMB_cv(
  data = catch.32603,
  total_catch_number ~ 1 + scale_doy2,
  offset = "log_effort",
  mesh = mesh,
  time = "sample_year",
  family = tweedie(link = "log"), 
  spatiotemporal = "IID",
  spatial = "on", 
  anisotropy = TRUE, 
  fold_ids = fold_vector
)

```

cross validation comparison
 - in general including spatial is better
```{r eval=FALSE, include=FALSE}
#1. compare the nge log likelihood - closer to 0 is better
fit2.cv$sum_loglik
fit1.cv$sum_loglik # way closer to 0

#2. look at mean log likelihoods
mean_loglike_fit2 <- fit2.cv$sum_loglik / nrow(fit2.cv$data) 
mean_loglike_fit1 <- fit1.cv$sum_loglik / nrow(fit1.cv$data)
#3. look at RMSE across entire dataset
sqrt(mean((fit2.cv$data$total_catch_number - fit2.cv$data$cv_predicted)^2)) 
sqrt(mean((fit1.cv$data$total_catch_number - fit1.cv$data$cv_predicted)^2)) 
#4. RMSE by folds
rmse_fit2 <- dplyr::group_by(fit2.cv$data, cv_fold) |> 
  dplyr::summarize(
    mae = mean(abs(total_catch_number - cv_predicted)),
    rmse = sqrt(mean((total_catch_number - cv_predicted)^2)))
rmse_fit1 <- dplyr::group_by(fit1.cv$data, cv_fold) |> 
  dplyr::summarize(
    mae = mean(abs(total_catch_number - cv_predicted)),
    rmse = sqrt(mean((total_catch_number - cv_predicted)^2)))
# hmm look the same
ggplot() +
  geom_line(aes(x = cv_fold, rmse), data = rmse_fit2, col = "red", alpha = 0.7, linewidth = 1.5) +
  geom_line(aes(x = cv_fold, rmse), data = rmse_fit1, col = "blue")

```



weight model fit
```{r}
fit.wgt <- sdmTMB(
  data = catch.32603,
  # we're curious about the total_catch_number (abundance) that may vary by time (so estimating a diff parameter for each year) 
  total_catch_weight ~ 0 + as.factor(sample_year) + scale_doy2,
  offset = log(catch.32603$effort),
  # this is the mesh we created above that (we think) is a finer scale mesh, can be coarser if conversion issues arise
  mesh = mesh,
  # since we have spatio temporal fields on (w/ iid) we need to indicate what our time column is
  time = "sample_year",
  # tweedie family is a good starting point generally used with continuous and/or zero inflated biomass data the
  # catch data is perhaps not the most appropriate but we'll start here, maybe poisson with a catch offset would be better?
  family = tweedie(link = "log"), 
  # spatiotmeporal field is iid meaning each year is independently estimated 
  # this is another spot we could loosen up on flexibility if it doesn't converge (i.e., ar1 or rw)
  spatiotemporal = "iid",
  spatial = "on", 
  # anisotropy v isotrophy this is the decorrelation assumptions with Matern distance. 
  # basically it assume that the decorrelation dist the same E-W and N-S however in the BS 
  # we would expect that there would be greater correlation along shelf (NW/SE) than across shelf E-W
  # in general keep unless its unestimable 
  anisotropy = TRUE
)
```



## model checks
```{r}
# passes the standard sanity checks
sanity(fit1); sanity(fit.wgt)
# things of interest here, 
#' dispersion - how much variability after account for main effects and spatial/spatiotemporal 
#' matern range - indicates if smooth or patchy
#' spatiial SD if high means theres strong peaks / valleys in spatial fields, if 0 no spatial variability
# how do the two spatial on and spatial off models do
tidy(fit1, "ran_pars"); tidy(fit2, "ran_pars")
# the spatial E paramter gets larger when spatial is off (duh)
tidy(fit.wgt, "ran_pars")
#' range smooth or patchy (but which is which)

# looking for approximately normal along the line
catch.32603$resid <- residuals(fit1, type = "mle-mvn")
qqnorm(catch.32603$resid)
abline(0,1)

hist(catch.32603$resid)

# can also look at the simulated residuals
sims <- simulate(fit1, nsim = 500, type = "mle-mvn")
resids <- dharma_residuals(sims, fit1, return_DHARMa = TRUE)
p_qq <- as.grob(~DHARMa::plotQQunif(resids))
p_res <- as.grob(~DHARMa::plotResiduals(resids))
  
# combine residual plots
dharma_plots <- wrap_plots(p_qq, p_res) +
plot_annotation(title = paste("DHARMa Diagnostics"))
#ggsave(plot = dharma_plots, filename = here("figures", "pcod_catch_tweedie_diagnostics.png"))

catch.32603$resid <- resids$scaledResiduals

# spatial residual plots
spatials <- ggplot(catch.32603) +
  geom_point(aes(X,Y, col = resid)) +
  labs(title = paste("Pcod catch tweedie Spatial residuals")) +
  scale_colour_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  facet_wrap(~sample_year) +
  coord_fixed() +
  theme_minimal()

#ggsave(plot = spatials, filename = here("figures", "pcod_catch_tweedie_spatial_diagnostics.png"))

# weight residuals
catch.32603$wgt_resid <- residuals(fit.wgt, type = "mle-mvn")
qqnorm(catch.32603$wgt_resid)
abline(0,1)

hist(catch.32603$wgt_resid)

# spatial residuals
# here we're revuing the data to make sure we're not seeing patterns of clustering of
# positive and negative residuals together
ggplot() +
  geom_point(data = catch.32603, aes(X, Y, color = wgt_resid)) +
  scale_color_gradient2() +
  facet_wrap(~sample_year)
```

## prediction grids
```{r}
library(sf)
library(sfheaders)
library(concaveman)
# steps : 
# create sf object, converts to metric
catch.sf <- catch.32603 %>%
  dplyr::select(c(X, Y, eq_latitude, eq_longitude)) %>%
  mutate(X = X * 1000, Y = Y * 1000) %>% # convert to meters for crs 32603
  st_as_sf(coords = c("X", "Y"), crs = 32603)

# check to make sure it makes sense
head(catch.sf$geometry)
# create convex hull 
convex_hull <- concaveman(catch.sf, concavity = 3)
plot(convex_hull)

# add buffer
convex_buff <- st_buffer(convex_hull, dist = 15e3) # dist in meters (i think) so this is 15 km buffer
plot(convex_buff)

# make grid, created a cell size of 10,000 x 10,000 m (10 x 10 km) with poitns in center
# need to check units on this... 
grid <- st_make_grid(convex_buff, cellsize = 10e3, what = "centers") %>%
  st_intersection(convex_buff)
plot(grid)

coarse.grid <- st_make_grid(convex_buff, cellsize = 40e3, what = "centers") %>%
  st_intersection(convex_buff)
plot(coarse.grid)

# change back to df, this wasn't straightforward for me...
bs.grid <- do.call(rbind, st_geometry(grid)) %>%
  as.tibble() %>% setNames(c("X", "Y")) %>%
  # convert df m back to km
  mutate(X = X/1e3, Y = Y/1e3)

plot(bs.grid)

coarse.grid.df <- do.call(rbind, st_geometry(coarse.grid)) %>%
  as.tibble() %>% setNames(c("X", "Y")) %>%
  # convert df m back to km
  mutate(X = X/1e3, Y = Y/1e3)

# replicate the df grid for each year (or other variables we include either spatio temporal or temporal) depth, temp etc
bs.grid.tmp <- replicate_df(bs.grid, "sample_year", unique(catch.32603$sample_year)) # does this across 2003:2025, but no 2020 because theres no survey then
coarse.grid.tmp <- replicate_df(coarse.grid.df, "sample_year", unique(catch.32603$sample_year))

scaled_median <- unique(catch.32603$scaled_median)
bs.grid.yrs <- replicate_df(bs.grid.tmp, "scale_doy2", scaled_median)
coarse.grid.yrs <- replicate_df(coarse.grid.tmp, "scale_doy2", scaled_median)
```

## predict across grid
```{r}
preds <- predict(fit1, type = "link", newdata = bs.grid.yrs, return_tmb_object = F) %>% # catch
  mutate(Y_m = Y * 1000,
         X_m = X * 1000) # sf requires meters, predicts on the link scale (log)
preds.wgt <- predict(fit.wgt, newdata = bs.grid.yrs, return_tmb_object = F) %>%  # weight
  mutate(Y_m = Y * 1000,
         X_m = X * 1000) # sf requires meters

preds.sf <- st_as_sf(preds, coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

preds.wgt.sf <- st_as_sf(preds.wgt, coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

# create sf of catch stations
catch.sf <- catch.32603 %>%
  mutate(X_m = X * 1000,
         Y_m = Y * 1000) %>%
  st_as_sf(coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

#preds.sf.red <- filter(preds.sf, est < 1e4)
# predict the catch across the grid, catch units (if in link or response space are based on the arg of the predict function above. )
num.spatiotemp.preds.map <- ggplot() +
  geom_sf(preds.sf, mapping = aes(fill = est, color = est), pch = 22) +
  scale_fill_viridis_c() + scale_color_viridis_c() +
  geom_sf(catch.sf, mapping = aes(shape = total_catch_number > 0)) +
  scale_shape_manual(values = c(4, 19)) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude", color = "Log Catch Number",
       fill = "Log Catch Number", shape = "More than 0 fish caught") +
  theme(text = element_text(size = 16))

# spatial field
ggplot() +
  geom_sf(preds.sf, mapping = aes(fill = omega_s, color = omega_s), pch = 22) +
  scale_fill_viridis_c() + scale_color_viridis_c()

# predict the catch weight across the grid, catch weight needs to be logged because of the 
# tweedie log link
wgt.spatiotemp.preds.map <- ggplot() +
  geom_sf(preds.wgt.sf, mapping = aes(fill = est, color = est), pch = 22) +
  scale_fill_viridis_c() + scale_color_viridis_c() +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_sf(catch.sf, mapping = aes(shape = total_catch_weight > 0)) +
  scale_shape_manual(values = c(4, 19)) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude", color = "Log Catch Weight",
       fill = "Log Catch Weight") +
  theme(text = element_text(size = 16))

# ggsave(filename = "figures/catch_num_sptemp_map_072726.png", plot = num.spatiotemp.preds.map,
#        height = 14, width = 14)
# ggsave(filename = "figures/catch_wgt_sptemp_map_051226.png", plot = wgt.spatiotemp.preds.map,
#        height = 14, width = 14)
```

## predict obs
```{r}
preds <- predict(fit1)

#' est: fixed + spatial + spatiotemporal (use this for maps and indices)
#' est_non_rf - fixed effects only, use this to understand the covariate-driven signal (in our case here this would
#' be the temporal effects)
#' omega_s - spatial random effects use this to see persistent spatial deviations 
#' epsilon_st - spatiotemporal random effects, will tell you/show year-specific anomalies


# i think this is the spatiotemporal random effects fields
preds %>% 
  ggplot(aes(x=X, y=Y, color=epsilon_st)) + 
  #theme_linedraw() +
  scale_color_viridis_c() +
  geom_point() +
  facet_wrap(~sample_year) +
  ggtitle("Spatiotemporal RE (link space)",subtitle="Spatiotemporal Model")

# to includes overall estimate that includes the main effects

preds %>%
  ggplot(aes(x =X, y = Y, color = est)) +
  scale_color_viridis_c() +
  geom_point() +
  facet_wrap(~sample_year)
  
```

## get indicies

```{r}
#index_preds <- predict(fit1, return_tmb_object = TRUE)
?get_cog
#get_index(index_preds)

# try a coarse grid for cog predictions
index_preds_coarse <- predict(fit1, newdata = coarse.grid.yrs, return_tmb_object = TRUE)
index_preds <- predict(fit1, newdata = bs.grid.yrs, return_tmb_object = TRUE)
# we don't specifically indicate the offset, but thats okay here because we're estimating
# relative biomass (since area =1 and we scaled it down from the grid size of 100 km2)
# which is the relative catch in numbers of the EBS (in the top 20m of the surface, no expansion for dpeth)
index <- get_index(index_preds, bias_correct = T, area = 1)
# this is catch numbers

# index_preds_wgt <- predict(fit.wgt, newdata = bs.grid.yrs, return_tmb_object = TRUE)
# index.wgt <- get_index(index_preds_wgt, bias_correct = T)

# visualize the index 
library(scales)
ggplot(data = index, aes(sample_year, est)) +
  geom_pointrange(aes(sample_year, ymin = lwr, ymax = upr)) +
  scale_y_continuous(labels = label_comma()) +
  labs(x = "Year", y = "Relative index of abundance (no.)")

# ggplot(index.wgt) +
#   geom_point(aes(sample_year, log(est))) +
#   geom_errorbar(aes(sample_year, ymin = lwr, ymax = upr))

# Note that get_cog with bias_correct = T runs out of RAM on the computer
# for initial exploration you can use the bias_correct = F
# for final interpretation bias_correct = T but you have likely include a nsplit argument 
# and split up the predictions and do it separately. You could also predict on a coarser
# grid, right now we use a grid of 10 x 10 km which is decent
cog <- get_cog(index_preds, bias_correct = F, format = "wide")
head(cog)

# trying a coarse grid to predict
cog2 <- get_cog(index_preds_coarse, bias_correct = T, format = "wide")
head(cog2) # sweet deal it didn't expode the computer. 


# quick plot
ggplot(cog2, aes(est_x, est_y, colour = sample_year)) +
  geom_pointrange(aes(xmin = lwr_x, xmax = upr_x)) +
  geom_pointrange(aes(ymin = lwr_y, ymax = upr_y)) +
  scale_colour_viridis_c()

cog.4326 <- cog2 %>%
  mutate(est_x_m = est_x*1000,
         est_y_m = est_y*1000) %>%
  st_as_sf(coords = c("est_x_m", "est_y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude() 

cog.df <- cog.4326 %>%
  mutate(est_y_4326 = st_coordinates(.)[,1],
         est_x_4326 = st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
  dplyr::select(c("sample_year", "est_y_4326", "est_x_4326"))

cog.error.min <- cog2 %>%
  mutate(lwr_x = lwr_x * 1000,
         lwr_y = lwr_y * 1000) %>%
  st_as_sf(coords = c("lwr_x", "lwr_y"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude() %>%
  mutate(lon_min = st_coordinates(.)[,1],
         lat_min = st_coordinates(.)[,2]) %>%
  select(sample_year, lon_min, lat_min) %>%
  st_drop_geometry()

cog.error.max <- cog2 %>%
  mutate(upr_x = upr_x * 1000,
         upr_y = upr_y * 1000) %>%
  st_as_sf(coords = c("upr_x", "upr_y"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude() %>%
  mutate(lon_max = st_coordinates(.)[,1],
         lat_max = st_coordinates(.)[,2]) %>%
  select(sample_year, lon_max, lat_max) %>%
  st_drop_geometry()

cog.error <- left_join(cog.error.min, cog.error.max, by = "sample_year") %>%
  left_join(cog.df)

ggplot(cog2) +
  geom_point(aes(est_x, est_y)) +
  ggrepel::geom_text_repel(aes(est_x, est_y, label = sample_year))
  

# put the cog on a map
cog_plot <- ggplot() +
  # geom_point(data = cog, aes(est_x, est_y)) +
  # geom_errorbar(data = cog, aes(x = est_x, y = est_y, xmin = lwr_x, xmax = upr_x)) +
  # geom_errorbar(data = cog, aes(x = est_x, y = est_y, ymin = lwr_y, ymax = upr_y))
  # 
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F)  +
  geom_sf(data = cog.4326, aes(color = sample_year)) +
  geom_errorbar(data = cog.error, aes(x = est_y_4326, ymin = lat_min, ymax = lat_max,
                                      color = sample_year)) +
  geom_errorbarh(data = cog.error, aes(y = est_x_4326, xmin = lon_min, xmax = lon_max,
                                       color = sample_year)) +
  coord_sf(crs = st_crs(4326)) +
  ggrepel::geom_label_repel(data = filter(cog.error, sample_year == 2007), 
                           aes(x = est_y_4326, y = est_x_4326, label = sample_year)) +
  scale_colour_viridis_c() +
  labs(x = "Longitude", y = "Latitude") +
  # geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
  #              breaks = c(-50, -70, -150), color = "black") +
  theme_minimal()

# ggsave(plot = cog_plot, filename = here("figures", "Pcod_abund_cog_4326.png"),
#        height = 10, width = 14)
```


