---
title: "3_catch_sdmTMB_exploration"
output: html_document
date: "2026-03-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#libraries
```{r}
library(EMAdownload)
library(tidyverse)
library(ggplot2)
library(sdmTMB)
library(patchwork)
library(sf)
library(marmap)
```

# data
```{r}
this.year <- format(Sys.Date(), "%Y") %>% as.numeric()

pcod_catch <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN", catch0 = TRUE) 
# note akfin upload date, if in the upload date says some time in Jan 2026 then the pcod data hasn't been updated to include the correct pcod cut off point for 140 mm

pcod_catch_nets <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "NETS156", catch0 = TRUE) 
```

# create base/bathy maps
```{r}
# make sure the base map / bathy is created with the advanced sf library 
sf_use_s2(TRUE)

# need to use world not world2hires cause that one is hella old and doesn't plot correctly
bs_map <- maps::map("world", regions = c("USA:Alaska", "Russia"), 
                    fill = TRUE, plot = FALSE, wrap = c(0, 360)) %>%
  st_as_sf(crs = 4326) 

# Use bounding box, have to add 360 to the bbox coords
# xmin = 185 (-175), xmax = 202 (-155)
bbox_360 <- st_bbox(c(xmin = 180, xmax = 210, ymin = 50, ymax = 68), 
                    crs = st_crs(bs_map))

# crop 
# Using st_make_valid helps w/ messy polys
bs_base <- bs_map %>% 
  st_make_valid() %>% 
  st_intersection(st_as_sfc(bbox_360)) %>%
  st_shift_longitude()

bs_croppedUTM <- bs_base %>%
  st_transform(crs = 32602) # Move to UTM Zone 2N

#read in bathymetry file.  
bathline <- getNOAA.bathy(-180,-155,53,68)

bathline.xyz <- as.xyz(bathline)
```


go through some of the basic steps for sdmTMB 

- add_utm_columns
- make mesh, make_mes()
- fit model 
    - modelling catch data with effort data
    - for now i think we're more interestd in catch data rather than weightcpue
    - an easy start for distribution would be to use the tweedie model (alt. could use a poisson with a effort offset)
    
```{r}
# check to make sure the format makes sense
head(pcod_catch)
unique(pcod_catch$common_name)
range(pcod_catch$total_catch_number)

table(pcod_catch$total_catch_number > 0) # lots of zeros 
table(pcod_catch$sample_year, pcod_catch$total_catch_number >0)

hist(pcod_catch$total_catch_number) # kinda looks zero inflated :(
# for now lets move forward with it the way it is and see how model fits look

table(pcod_catch$lhs_code, pcod_catch$total_catch_number > 0)

pcod_catch_a0 <- filter(pcod_catch, lhs_code == "A0")

filter(pcod_catch_a0, is.na(effort)) # mising effort at DY1408, station 20140308022

test <-filter(pcod_catch_a0, total_catch_weight < 0)

sample.size <-pcod_catch_a0 %>%
  group_by(sample_year, region) %>%
  dplyr::summarise(total_catch = sum(total_catch_number),
                   total_weight = sum(total_catch_weight))

pcod_catch_a0 %>%
  group_by(region) %>%
  dplyr::summarise(mean_catch = mean(total_catch_number))
```

## add utm columns
```{r}
catch.32603 <- pcod_catch_a0 %>%
  filter(!is.na(effort)) %>% # remove instance of missing effort 20140308022
  filter(total_catch_weight != -999) %>% # remove weird catch value, that should be NA
  add_utm_columns(ll_names = c("eq_longitude", "eq_latitude"),
                ll_crs = 4326,
                units = "km",
                utm_names = c("X", "Y")) %>%
  dplyr::select(c(sample_year, station_id, master_station_name:region, 
                  eq_time:eq_longitude, effort, effort_units, 
                  species_tsn:total_catch_weight, X, Y)) %>%
  mutate(cpue = total_catch_number / effort,
         total_catch_weight = total_catch_weight / 1000)
# this uses CRS 32603
```

## make mesh
```{r}
# this is somewhat of a high number of knots (I think) but lets try it out for now
mesh <- make_mesh(catch.32603, 
                  c("X", "Y"), 
                  n_knots = 80)
plot(mesh)
# might want to consider some kind of barrier here to make sure that we're not trying to make a correlation matrix thingy with places with no data that don't make sense
```

## fit model

this is a first pass to see if this converges but these are why i think i'm making these decisions
```{r}
# this model estimates: independent year effects 
# (because of sample_year in model formula),
fit1 <- sdmTMB(
  data = catch.32603,
  # we're curious about the total_catch_number (abundance) that may vary by time (so estimating a diff parameter for each year) 
  total_catch_number ~ 1 + as.factor(sample_year),
  offset = log(catch.32603$effort),
  # this is the mesh we created above that (we think) is a finer scale mesh, can be coarser if conversion issues arise
  mesh = mesh,
  # since we have spatio temporal fields on (w/ iid) we need to indicate what our time column is
  time = "sample_year",
  # tweedie family is a good starting point generally used with continuous and/or zero inflated biomass data the
  # catch data is perhaps not the most appropriate but we'll start here, maybe poisson with a catch offset would be better?
  family = tweedie(link = "log"), 
  # spatiotmeporal field is iid meaning each year is independently estimated 
  # this is another spot we could loosen up on flexibility if it doesn't converge (i.e., ar1 or rw)
  spatiotemporal = "IID",
  # spatial is on when spatiotemporal is off, but it can be on with spatiotemporal
  spatial = "off", 
  # anisotropy v isotrophy this is the decorrelation assumptions with Matern distance. 
  # basically it assume that the decorrelation dist the same E-W and N-S however in the BS 
  # we would expect that there would be greater correlation along shelf (NW/SE) than across shelf E-W
  # in general keep unless its unestimable 
  anisotropy = TRUE
)
tidy(fit1, "ran_pars")
```

weight model fit
```{r}
fit.wgt <- sdmTMB(
  data = catch.32603,
  # we're curious about the total_catch_number (abundance) that may vary by time (so estimating a diff parameter for each year) 
  total_catch_weight ~ 0 + as.factor(sample_year),
  offset = log(catch.32603$effort),
  # this is the mesh we created above that (we think) is a finer scale mesh, can be coarser if conversion issues arise
  mesh = mesh,
  # since we have spatio temporal fields on (w/ iid) we need to indicate what our time column is
  time = "sample_year",
  # tweedie family is a good starting point generally used with continuous and/or zero inflated biomass data the
  # catch data is perhaps not the most appropriate but we'll start here, maybe poisson with a catch offset would be better?
  family = tweedie(link = "log"), 
  # spatiotmeporal field is iid meaning each year is independently estimated 
  # this is another spot we could loosen up on flexibility if it doesn't converge (i.e., ar1 or rw)
  spatiotemporal = "IID",
  spatial = "off", 
  # anisotropy v isotrophy this is the decorrelation assumptions with Matern distance. 
  # basically it assume that the decorrelation dist the same E-W and N-S however in the BS 
  # we would expect that there would be greater correlation along shelf (NW/SE) than across shelf E-W
  # in general keep unless its unestimable 
  anisotropy = TRUE
)
```



## model checks
```{r}
# passes the standard sanity checks
sanity(fit1); sanity(fit.wgt)
# things of interest here, 
#' dispersion - how much variability after account for main effects and spatial/spatiotemporal 
#' matern range - indicates if smooth or patchy
#' spatiial SD if high means theres strong peaks / valleys in spatial fields, if 0 no spatial variability
#' 
tidy(fit1, "ran_pars"); tidy(fit.wgt, "ran_pars")
#' range smooth or patchy (but which is which)

# looking for approximately normal along the line
catch.32603$resid <- residuals(fit1, type = "mle-mvn")
qqnorm(catch.32603$resid)
abline(0,1)

hist(catch.32603$resid)

# weight residuals
catch.32603$wgt_resid <- residuals(fit.wgt, type = "mle-mvn")
qqnorm(catch.32603$wgt_resid)
abline(0,1)

hist(catch.32603$wgt_resid)

# spatial residuals
# here we're revuing the data to make sure we're not seeing patterns of clustering of
# positive and negative residuals together
ggplot() +
  geom_point(data = catch.32603, aes(X, Y, color = resid)) +
  scale_color_gradient2() +
  facet_wrap(~sample_year)
# kinda looks okay?

ggplot() +
  geom_point(data = catch.32603, aes(X, Y, color = wgt_resid)) +
  scale_color_gradient2() +
  facet_wrap(~sample_year)
```

## prediction grids
```{r}
library(sf)
library(sfheaders)
library(concaveman)
# steps : 
# create sf object, converts to metric
catch.sf <- catch.32603 %>%
  dplyr::select(c(X, Y, eq_latitude, eq_longitude)) %>%
  mutate(X = X * 1000, Y = Y * 1000) %>% # convert to meters for crs 32606 (but we're using 32603 i think thats ok)
  st_as_sf(coords = c("X", "Y"), crs = 32603)

# check to make sure it makes sense
head(catch.sf$geometry)
# create convex hull 
convex_hull <- concaveman(catch.sf, concavity = 3)
plot(convex_hull)

# add buffer
convex_buff <- st_buffer(convex_hull, dist = 15e3) # dist in meters (i think) so this is 15 km buffer
plot(convex_buff)

# make grid, created a cell size of 10,000 x 10,000 m (10 x 10 km) with poitns in center
# need to check units on this... 
grid <- st_make_grid(convex_buff, cellsize = 10e3, what = "centers") %>%
  st_intersection(convex_buff)
plot(grid)

# change back to df, this wasn't straightforward for me...
bs.grid <- do.call(rbind, st_geometry(grid)) %>%
  as.tibble() %>% setNames(c("X", "Y")) %>%
  # convert df m back to km
  mutate(X = X/1e3, Y = Y/1e3)

plot(bs.grid)

# replicate the df grid for each year (or other variables we include either spatio temporal or temporal) depth, temp etc
bs.grid.yrs <- replicate_df(bs.grid, "sample_year", unique(catch.32603$sample_year)) # does this across 2003:2025, but no 2020

```

## predict across grid
```{r}
preds <- predict(fit1, newdata = bs.grid.yrs, return_tmb_object = F) %>% # catch
  mutate(Y_m = Y * 1000,
         X_m = X * 1000) # sf requires meters
preds.wgt <- predict(fit.wgt, newdata = bs.grid.yrs, return_tmb_object = F) %>%  # weight
  mutate(Y_m = Y * 1000,
         X_m = X * 1000) # sf requires meters

preds.sf <- st_as_sf(preds, coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

preds.wgt.sf <- st_as_sf(preds.wgt, coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

# create sf of catch stations
catch.sf <- catch.32603 %>%
  mutate(X_m = X * 1000,
         Y_m = Y * 1000) %>%
  st_as_sf(coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

# predict the catch across the grid, catch needs to be logged because of tweedie log link
num.spatiotemp.preds.map <- ggplot() +
  geom_sf(preds.sf, mapping = aes(fill = est, color = est), pch = 22) +
  scale_fill_viridis_c() + scale_color_viridis_c() +
  geom_sf(catch.sf, mapping = aes(shape = total_catch_number > 0)) +
  scale_shape_manual(values = c(4, 19)) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude", color = "Log Catch Number",
       fill = "Log Catch Number") +
  theme(text = element_text(size = 16))

# predict the catch weight across the grid, catch weight needs to be logged because of the 
# tweedie log link
wgt.spatiotemp.preds.map <- ggplot() +
  geom_sf(preds.wgt.sf, mapping = aes(fill = est, color = est), pch = 22) +
  scale_fill_viridis_c() + scale_color_viridis_c() +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_sf(catch.sf, mapping = aes(shape = total_catch_weight > 0)) +
  scale_shape_manual(values = c(4, 19)) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude", color = "Log Catch Weight",
       fill = "Log Catch Weight") +
  theme(text = element_text(size = 16))

# ggsave(filename = "figures/catch_num_sptemp_map_051226.png", plot = num.spatiotemp.preds.map,
#        height = 14, width = 14)
# ggsave(filename = "figures/catch_wgt_sptemp_map_051226.png", plot = wgt.spatiotemp.preds.map,
#        height = 14, width = 14)
```

## predict obs
```{r}
preds <- predict(fit1)

#' est: fixed + spatial + spatiotemporal (use this for maps and indices)
#' est_non_rf - fixed effects only, use this to understand the covariate-driven signal (in our case here this would
#' be the temporal effects)
#' omega_s - spatial random effects use this to see persistent spatial deviations 
#' epsilon_st - spatiotemporal random effects, will tell you/show year-specific anomalies


# i think this is the spatiotemporal random effects fields
preds %>% 
  ggplot(aes(x=X, y=Y, color=epsilon_st)) + 
  #theme_linedraw() +
  scale_color_viridis_c() +
  geom_point() +
  facet_wrap(~sample_year) +
  ggtitle("Spatiotemporal RE (link space)",subtitle="Spatiotemporal Model")

# to includes overall estimate that includes the main effects

preds %>%
  ggplot(aes(x =X, y = Y, color = est)) +
  scale_color_viridis_c() +
  geom_point() +
  facet_wrap(~sample_year)
  
```

## get indicies
jk we're going to wait on this cause we want to make a prediction grid first
```{r}
#index_preds <- predict(fit1, return_tmb_object = TRUE)
?get_cog
#get_index(index_preds)

index_preds <- predict(fit1, newdata = bs.grid.yrs, return_tmb_object = TRUE)
index <- get_index(index_preds, bias_correct = T)
# this is catch numbers

index_preds_wgt <- predict(fit.wgt, newdata = bs.grid.yrs, return_tmb_object = TRUE)
index.wgt <- get_index(index_preds_wgt, bias_correct = T)


ggplot(index) +
  geom_point(aes(sample_year, log(est))) +
  geom_errorbar(aes(sample_year, ymin = lwr, ymax = upr))

ggplot(index.wgt) +
  geom_point(aes(sample_year, log(est))) +
  geom_errorbar(aes(sample_year, ymin = lwr, ymax = upr))

cog <- get_cog(index_preds)
head(cog)

ggplot(cog) +
  geom_point(aes(sample_year, est)) +
  geom_errorbar(aes(sample_year, ymin = lwr, ymax = upr)) +
  facet_wrap(~coord, scales = "free_y")
```


