---
title: "4_lng_sdmTMB_expoloration"
output: html_document
date: "2026-05-04"
editor_options: 
  chunk_output_type: console
---

# library
```{r}
library(tidyverse)
library(EMAdownload)
library(sdmTMB)
library(lubridate)
library(sf)
library(concaveman)
library(maps)
library(marmap)
library(visreg)
```


# data
```{r}
# pre curated length data (see 1_extract_lng_data.Rmd)
pcod_lngs <- read.csv("data/pcod_length_corrected_050426.csv") %>% filter(length < 154) %>% dplyr::select(-X) %>%
  mutate(doy = yday(haul_date), # convert to doy and then scale to mean 0 sd 1
         scale_doy = scale(doy, center = TRUE, scale = TRUE))
this.year <- format(Sys.Date(), "%Y") %>% as.numeric()
pcod_catch <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN", catch0 = TRUE, lhs = "A0") 

```

basic checks
- what is the proportion of sampled pcod to the total unsampled (for lengths) catch numbers?
```{r}
# all NAs removed 
table(is.na(pcod_lngs$corrected_lng)) # okay lengths look good
#na.lngs <- filter(pcod_lngs, is.na(corrected_lng))

pcod_lng_size <- pcod_lngs %>%
  group_by(station_id) %>%
  dplyr::summarise(num_lng = n()) # 451 stations sampled

catch_prop <- left_join(pcod_catch, pcod_lng_size, by = "station_id") %>%
  group_by(station_id) %>%
  mutate(proportion_sampled = num_lng / total_catch_number)

range(catch_prop$proportion_sampled, na.rm = T)

hist(catch_prop$proportion_sampled)

# none with a proportion greater than 1 which is good
filter(catch_prop, proportion_sampled > 1)


# clean up the pcod length dataframe
lngs <- dplyr::select(pcod_lngs, c(sample_year, eq_latitude, eq_longitude, scientific_name,
                           lhs_code, species_tsn, specimen_notes, corrected_lng, doy, scale_doy))
```
If we wanted to feel decent about our proportion sampled, we could resample/bootstrap assignment of 
lengths to the unmeasured catch and look at the variability in the mean estimates. We did this with the pollock
data and found that variability went down if approx 10% or mroe of the catch was sampled

basic steps for sdmtmb
- add_utm_columns
- make mesh, make_mesh()
- fit model 
    - modelling length data (likely gaussian)
- create prediction grid
- predict & map predictions (both average spatial field but also annual estimates)


other steps
# create base/bathy map
```{r}
# make sure the base map / bathy is created with the advanced sf library 
sf_use_s2(TRUE)

# need to use world not world2hires cause that one is hella old and doesn't plot correctly
bs_map <- maps::map("world", regions = c("USA:Alaska", "Russia"), 
                    fill = TRUE, plot = FALSE, wrap = c(0, 360)) %>%
  st_as_sf(crs = 4326) 

# Use bounding box, have to add 360 to the bbox coords
# xmin = 185 (-175), xmax = 202 (-155)
bbox_360 <- st_bbox(c(xmin = 180, xmax = 210, ymin = 50, ymax = 68), 
                    crs = st_crs(bs_map))

# crop 
# Using st_make_valid helps w/ messy polys
bs_base <- bs_map %>% 
  st_make_valid() %>% 
  st_intersection(st_as_sfc(bbox_360)) %>%
  st_shift_longitude()

bs_croppedUTM <- bs_base %>%
  st_transform(crs = 32602) # Move to UTM Zone 2N

#read in bathymetry file.  
bathline <- getNOAA.bathy(-180,-155,53,68)

bathline.xyz <- as.xyz(bathline)
```

plot raw data
```{r}
pcod_lngs_sf <- pcod_lngs %>%
  st_as_sf(coords = c("eq_longitude", "eq_latitude"), crs = 4326) %>%
  st_shift_longitude()

ggplot() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F)
```


# sdmtmb length
## add_utm_columns
```{r}
pcod_lngs_utm <- add_utm_columns(lngs, 
                  ll_names = c("eq_longitude", "eq_latitude"),
                  utm_names = c("X", "Y"),
                  ll_crs = 4326,
                  units = "km") # units should be in km, but sf generally requires m (fyi)
# uses 32603 UTM Zone 3N
plot(pcod_lngs_utm$X, pcod_lngs_utm$Y)
```

## make mesh
```{r}
mesh <- make_mesh(pcod_lngs_utm,
                  xy_cols = c("X", "Y"),
                  n_knots = 60)
```

## fit model 
```{r}
fit <- sdmTMB(data = pcod_lngs_utm,
              # removes the intercept (0)
              log(corrected_lng) ~ 0 + as.factor(sample_year) + s(scale_doy),
              mesh = mesh,
              # needs to be here if you're estimating a spatiotemporal field
              time = "sample_year",
              # could also try lognormal(link = "log")
              family = gaussian(link = "identity"),
              spatial = "on",
              # each year independently estimated
              spatiotemporal = "IID",
              # we want this but if it doesn't fit we can try removing
              anisotropy = TRUE
              )
```

model checks
```{r}
sanity(fit)
tidy(fit, "ran_pars")
summary(fit)


## check residuals
pcod_lngs_utm$resid <- residuals(fit, type = "mle-mvn")
qqnorm(pcod_lngs_utm$resid)
abline(0,1)
# making sure that we're not seeing patterns in the residuals
ggplot() +
  geom_point(data = pcod_lngs_utm, aes(X, Y, color = resid)) +
  scale_color_gradient2() +
  facet_wrap(~sample_year)
# looks decent maybe some patterning in 2005?
```


## create prediction grid
```{r}
# create sf object
pcod.sf <- pcod_lngs_utm %>%
  ungroup() %>%
  dplyr::select(c(X, Y, eq_longitude, eq_latitude)) %>%
  mutate(X = X*1000, Y = Y*1000) %>% # needs to be converted to meters for sf to work
  st_as_sf(coords = c("X", "Y"), crs = 32603)

convex <- concaveman(pcod.sf, concavity = 3)
#plot(convex)
convex.buff <- st_buffer(convex, dist = 15e3)
#plot(convex.buff)

## make grid
# make grid, created a cell size of 10,000 x 10,000 m (10 x 10 km) with poitns in center
grid <- st_make_grid(convex.buff, cellsize = 10e3, what = "centers") %>%
  st_intersection(convex.buff)

## change back to df
bs.grid <- do.call(rbind, st_geometry(grid)) %>%
  as_tibble() %>% setNames(c("X", "Y")) %>%
  # convert df m back to km
  mutate(X = X/1000, Y = Y/1000)
plot(bs.grid)

## replicate df across years and include any other variables we inlcude in our fit of the model
## i.e., depth tmeperature etc
bs.grid.year <- replicate_df(bs.grid, "sample_year", unique(pcod_lngs$sample_year))
bs.grid.doy <- replicate_df(bs.grid.year, "scale_doy", 0) # between 235 and 236 is ~0


# maps some of the sites
# ggplot() +
#   geom_sf(data = pcod.sf.shift, aes()) +
#   geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
#   geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
#                breaks = c(-50, -70, -150), color = "black") +
#   #use the 360 coords for longitude (so -150 + 360)
#   coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F)
```

## predict & map predictions 
(both average spatial field but also annual estimates)

```{r}
# predictiosn would be good for fitting the average spatial field 
predictions <- predict(fit, newdata = bs.grid.doy) %>%
  mutate(Y_m = Y * 1000,
         X_m = X * 1000) # sf requires meters

predictions.sf <- st_as_sf(predictions, coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

ggplot(predictions, aes(X_m, Y_m, fill = exp(est))) +
  scale_fill_viridis_c() + geom_tile() +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude")

# needs to be exp(est) because you logtransformed your corrected length
lng.by.year <- ggplot() +
  geom_sf(data = predictions.sf, aes(fill = exp(est), color = exp(est)), pch = 22) +
  scale_color_viridis_c() + scale_fill_viridis_c() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude", color = "Predicted length (mm)",
       fill = "Predicted length (mm)") +
  theme(text = element_text(size = 16))

# average spatial field
lng.avg.spatial <- ggplot() +
  geom_sf(data = predictions.sf, aes(fill = exp(omega_s), color = exp(omega_s)), pch = 22) +
  scale_color_viridis_c() + scale_fill_viridis_c() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  labs(x = "Longitude", y = "Latitude", color = "Predicted length (mm)",
       fill = "Predicted length (mm)") +
  theme(text = element_text(size = 16))


# ggsave(filename = "figures/predicted_lng_year_050626.png", plot = lng.by.year,
#        height = 14, width = 14)
# 
# ggsave(filename = "figures/pred_lng_avg_spatial_052626.png", plot = lng.avg.spatial,
#        height = 14, width = 14)
```


## investigate avg length by year
```{r}
predictions_tmb <- predict(fit, newdata = bs.grid.doy, return_tmb_object = T)

visreg(fit, plot = F)
```

