---
title: "4_lng_sdmTMB_expoloration"
output: html_document
date: "2026-05-04"
editor_options: 
  chunk_output_type: console
---

# library
```{r}
library(tidyverse)
library(EMAdownload)
library(sdmTMB)
library(lubridate)
library(sf)
library(concaveman)
library(here)
library(maps)
library(marmap)
library(visreg)
```


# data
```{r}
# pre curated length data (see 1_extract_lng_data.Rmd)
pcod_lngs <- read.csv("data/pcod_length_corrected_050426.csv") %>% 
  # make sure the lngs are only ones below which we are sure are age 0 
  filter(length < 154) %>% dplyr::select(-X) %>%
  mutate(doy = yday(haul_date),
         median_doy = median(doy),
         scale_doy = doy - median_doy,
         mean_doy = mean(doy),
         sd_doy = sd(doy),
         scale_doy2 = (doy - mean_doy)/sd_doy,
         scaled_median = (median_doy - mean_doy) / sd_doy)


this.year <- format(Sys.Date(), "%Y") %>% as.numeric()


pcod_catch <- join_event_catch(start_year = 2003, end_year = this.year, survey_region = c("NBS", "SEBS"), tsn = 164711,
                trawl_method = "S", gear = "CAN", catch0 = TRUE, lhs = "A0") 

```

basic checks
- what is the proportion of sampled pcod to the total unsampled (for lengths) catch numbers?
```{r}
# all NAs removed 
table(is.na(pcod_lngs$corrected_lng)) # okay lengths look good
#na.lngs <- filter(pcod_lngs, is.na(corrected_lng))

pcod_lng_size <- pcod_lngs %>%
  group_by(station_id) %>%
  dplyr::summarise(num_lng = n()) # 451 stations sampled

nrow(pcod_lngs)


catch_prop <- left_join(pcod_catch, pcod_lng_size, by = "station_id") %>%
  group_by(station_id) %>%
  mutate(proportion_sampled = num_lng / total_catch_number)

range(catch_prop$proportion_sampled, na.rm = T)

hist(catch_prop$proportion_sampled)

# none with a proportion greater than 1 which is good
filter(catch_prop, proportion_sampled > 1)


pcod_lngs %>%
  group_by(region) %>%
  dplyr::summarise(num = n(),
                   mean = mean(n()))

# clean up the pcod length dataframe
lngs <- dplyr::select(pcod_lngs, c(sample_year, eq_latitude, eq_longitude, region, scientific_name,
                           lhs_code, species_tsn, specimen_notes, corrected_lng, doy, scale_doy,
                           scale_doy2, scaled_median))
```
If we wanted to feel decent about our proportion sampled, we could resample/bootstrap assignment of 
lengths to the unmeasured catch and look at the variability in the mean estimates. We did this with the pollock
data and found that variability went down if approx 10% or mroe of the catch was sampled

basic steps for sdmtmb
- add_utm_columns
- make mesh, make_mesh()
- fit model 
    - modelling length data (likely gaussian)
- create prediction grid
- predict & map predictions (both average spatial field but also annual estimates)


other steps
# create base/bathy map
```{r}
# make sure the base map / bathy is created with the advanced sf library 
sf_use_s2(TRUE)

# need to use world not world2hires cause that one is hella old and doesn't plot correctly
bs_map <- maps::map("world", regions = c("USA:Alaska", "Russia"), 
                    fill = TRUE, plot = FALSE, wrap = c(0, 360)) %>%
  st_as_sf(crs = 4326) 

# Use bounding box, have to add 360 to the bbox coords
# xmin = 185 (-175), xmax = 202 (-155)
bbox_360 <- st_bbox(c(xmin = 180, xmax = 210, ymin = 50, ymax = 68), 
                    crs = st_crs(bs_map))

# crop 
# Using st_make_valid helps w/ messy polys
bs_base <- bs_map %>% 
  st_make_valid() %>% 
  st_intersection(st_as_sfc(bbox_360)) %>%
  st_shift_longitude()

bs_croppedUTM <- bs_base %>%
  st_transform(crs = 32602) # Move to UTM Zone 2N

#read in bathymetry file.  
bathline <- getNOAA.bathy(-180,-155,53,68)

bathline.xyz <- as.xyz(bathline)
```

plot raw data
```{r}
pcod_lngs_sf <- lngs %>%
  filter(!(sample_year == 2007 & eq_latitude > 61.6)) %>%
  st_as_sf(coords = c("eq_longitude", "eq_latitude"), crs = 4326) %>%
  st_shift_longitude()

ggplot() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F)

ggplot() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~sample_year)

lng.2023 <- filter(pcod_lngs_sf, sample_year == 2023)
```


# sdmtmb length
## add_utm_columns
```{r}
pcod_lngs_utm <- add_utm_columns(lngs, 
                  ll_names = c("eq_longitude", "eq_latitude"),
                  utm_names = c("X", "Y"),
                  ll_crs = 4326,
                  units = "km") %>% # units should be in km, but sf generally requires m (fyi)
# uses 32603 UTM Zone 3N
  mutate(sample_year = as.factor(sample_year)) %>%
# what if we removed the fish caught in norton sound (maybe all of nbs) in 2007
  filter(!(sample_year == 2007 & eq_latitude > 61.6)) # to include the few fish caught at 61.51,
# this drops 200 lngs

plot(pcod_lngs_utm$X, pcod_lngs_utm$Y)
```

## make mesh
```{r}
mesh <- make_mesh(pcod_lngs_utm,
                  xy_cols = c("X", "Y"),
                  n_knots = 100)
```

## fit model 
```{r}
fit <- sdmTMB(data = pcod_lngs_utm,
              # removes the intercept (0)
              # added a spline to allow non linear variability of length with doy
              corrected_lng ~ 0 + as.factor(sample_year) + s(scale_doy2),
              mesh = mesh,
              # needs to be here if you're estimating a spatiotemporal field
              time = "sample_year",
              # could also try lognormal(link = "log")
              # trying also gaussian with log link
              family = gaussian(link = "log"),
              spatial = "on",
              # each year independently estimated
              spatiotemporal = "IID",
              # we want this but if it doesn't fit we can try removing
              anisotropy = TRUE
              )

fit.log <- sdmTMB(data = pcod_lngs_utm,
              # removes the intercept (0)
              # added a spline to allow non linear variability of length with doy
              corrected_lng ~ 0 + as.factor(sample_year) + s(scale_doy2),
              mesh = mesh,
              # needs to be here if you're estimating a spatiotemporal field
              time = "sample_year",
              family = lognormal(link = "log"),
              spatial = "on",
              # each year independently estimated
              spatiotemporal = "IID",
              # we want this but if it doesn't fit we can try removing
              anisotropy = TRUE
              )

fit.log2 <- sdmTMB(data = pcod_lngs_utm,
              # removes the intercept (0)
              # added a spline to allow non linear variability of length with doy
              corrected_lng ~ 0 + as.factor(sample_year) + s(scale_doy2),
              mesh = mesh,
              # needs to be here if you're estimating a spatiotemporal field
              time = "sample_year",
              family = lognormal(link = "log"),
              spatial = "off",
              # each year independently estimated
              spatiotemporal = "IID",
              # we want this but if it doesn't fit we can try removing
              anisotropy = TRUE
              )
```

### model evaluation
```{r}
sanity(fit)
sanity(fit.log)
sanity(fit.log2)
# initial look both models have converged
tidy(fit, "ran_pars")
tidy(fit.log, "ran_pars") # variability in the phi parameters, smaller sd for phi, sigma o and e
tidy(fit.log2, "ran_pars")
summary(fit)

AIC(fit); AIC(fit.log); AIC(fit.log2)
# in general the lognormal looks better than the gaussian
# the residuals look somewhat similar but AIC def agrees
# including the spatial is better - best way to evaluat is cv


## check residuals
pcod_lngs_utm$resid <- residuals(fit, type = "mle-mvn")
qqnorm(pcod_lngs_utm$resid)
abline(0,1)
# making sure that we're not seeing patterns in the residuals
ggplot() +
  geom_point(data = pcod_lngs_utm, aes(X, Y, color = resid)) +
  scale_color_gradient2() +
  facet_wrap(~sample_year)
# looks decent maybe some patterning in 2005?

## use simulated residuals
# can also look at the simulated residuals
sims <- simulate(fit, nsim = 500, type = "mle-mvn")
resids <- dharma_residuals(sims, fit, return_DHARMa = TRUE)
p_qq <- as.grob(~DHARMa::plotQQunif(resids))
p_res <- as.grob(~DHARMa::plotResiduals(resids))
  
# combine residual plots
dharma_plots <- wrap_plots(p_qq, p_res) +
plot_annotation(title = paste("Length gaussian without 2007 N. of 61.6 dist DHARMa Diagnostics"))
#ggsave(plot = dharma_plots, filename = here("figures", "pcod_red_lng_gaussianlog_diagnostics.png"))

# spatial residuals
pcod_lngs_utm$scl_resid <- resids$scaledResiduals

# spatial residual plots
spatials <- ggplot(pcod_lngs_utm) +
  geom_point(aes(X,Y, col = scl_resid)) +
  labs(title = paste("Pcod lng gaussian without 2007 N. of 61.6 Spatial residuals")) +
  scale_colour_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  facet_wrap(~sample_year) +
  coord_fixed() +
  theme_minimal()

#ggsave(plot = spatials, filename = here("figures", "pcod_red_lng_gaussian_spatial_diagnostics.png"))



### repeat with the lognormal 
sims.log <- simulate(fit.log, nsim = 500, type = "mle-mvn")
resids2 <- dharma_residuals(sims.log, fit.log, return_DHARMa = TRUE)
p_qq2 <- as.grob(~DHARMa::plotQQunif(resids2))
p_res2 <- as.grob(~DHARMa::plotResiduals(resids2))
  
# combine residual plots
dharma_plots2 <- wrap_plots(p_qq2, p_res2) +
plot_annotation(title = paste("Length lognormal without 2007 N. of 61.6 dist DHARMa Diagnostics"))
# ggsave(plot = dharma_plots2, filename = here("figures", "pcod_lng_lognormal_diagnostics.png"))

# spatial residuals
pcod_lngs_utm$scl_resid2 <- resids2$scaledResiduals

# spatial residual plots
spatials2 <- ggplot(pcod_lngs_utm) +
  geom_point(aes(X,Y, col = scl_resid2)) +
  labs(title = paste("Pcod lng lognormal Spatial residuals")) +
  scale_colour_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  facet_wrap(~sample_year) +
  coord_fixed() +
  theme_minimal()

# ggsave(plot = spatials2, filename = here("figures", "pcod_lng_logrnormal_spatial_diagnostics.png"))

```

### cross validation

dont run this unless you really want to - its takes forever and sometimes doesn't converge if you're using the reduced model (no norton sound data in 2007)
```{r eval=FALSE, include=FALSE}
fold_vector <- as.numeric(as.factor(pcod_lngs_utm$sample_year))

# open the parallel processing 
TMB::openmp(n = 1)
# set up paralell processing
future::plan(future::multisession, workers = 4)

fit.cv <- sdmTMB_cv(data = pcod_lngs_utm,
              corrected_lng ~ 0 + s(scale_doy2),
              mesh = mesh,
              time = "sample_year",
              family = gaussian(link = "log"),
              spatial = "on",
              spatiotemporal = "IID",
              anisotropy = TRUE,
              fold_ids = fold_vector
              )

fit.log.cv <- sdmTMB_cv(data = pcod_lngs_utm,
              corrected_lng ~ 0 + s(scale_doy2),
              mesh = mesh,
              time = "sample_year",
              family = lognormal(link = "log"),
              spatial = "on",
              spatiotemporal = "IID",
              anisotropy = TRUE,
              fold_ids = fold_vector
              )

fit.log2.cv <- sdmTMB_cv(data = pcod_lngs_utm,
              corrected_lng ~ 0 + s(scale_doy2),
              mesh = mesh,
              time = "sample_year",
              family = lognormal(link = "log"),
              spatial = "off",
              spatiotemporal = "IID",
              anisotropy = TRUE,
              fold_ids = fold_vector
              )
```

```{r}
#1. compare the nge log likelihood - closer to 0 is better
fit.cv$sum_loglik
fit.log.cv$sum_loglik 
fit.log2.cv$sum_loglik

#2. look at mean log likelihoods
mean_loglike_fit <- fit.cv$sum_loglik / nrow(fit.cv$data) 
mean_loglike_fit.log <- fit.log.cv$sum_loglik / nrow(fit.log.cv$data)
mean_loglike_fit.log2 <- fit.log2.cv$sum_loglik / nrow(fit.log2.cv$data)
#3. look at RMSE across entire dataset
sqrt(mean((fit.cv$data$total_catch_number - fit.cv$data$cv_predicted)^2)) 
sqrt(mean((fit.log.cv$data$total_catch_number - fit.log.cv$data$cv_predicted)^2)) 
sqrt(mean((fit.log2.cv$data$total_catch_number - fit.log2.cv$data$cv_predicted)^2)) 
#4. RMSE by folds
rmse_fit <- dplyr::group_by(fit.cv$data, cv_fold) |> 
  dplyr::summarize(
    mae = mean(abs(corrected_lng - cv_predicted)),
    rmse = sqrt(mean((corrected_lng - cv_predicted)^2)))
rmse_fit.log <- dplyr::group_by(fit.log.cv$data, cv_fold) |> 
  dplyr::summarize(
    mae = mean(abs(corrected_lng - cv_predicted)),
    rmse = sqrt(mean((corrected_lng - cv_predicted)^2)))
rmse_fit.log2 <- dplyr::group_by(fit.log2.cv$data, cv_fold) |> 
  dplyr::summarize(
    mae = mean(abs(corrected_lng - cv_predicted)),
    rmse = sqrt(mean((corrected_lng - cv_predicted)^2)))
# hmm look the same
ggplot() +
  geom_line(aes(x = cv_fold, rmse), data = rmse_fit, col = "red") +
  geom_line(aes(x = cv_fold, rmse), data = rmse_fit.log, col = "blue") +
  geom_line(aes(x = cv_fold, rmse), data = rmse_fit.log2, col = "green")
```


## create prediction grid
```{r}
# create sf object
pcod.sf <- pcod_lngs_utm %>%
  ungroup() %>%
  dplyr::select(c(X, Y, eq_longitude, eq_latitude)) %>%
  mutate(X = X*1000, Y = Y*1000) %>% # needs to be converted to meters for sf to work
  st_as_sf(coords = c("X", "Y"), crs = 32603)

convex <- concaveman(pcod.sf, concavity = 3)
#plot(convex)
convex.buff <- st_buffer(convex, dist = 15e3)
#plot(convex.buff)

## make grid
# make grid, created a cell size of 10,000 x 10,000 m (10 x 10 km) with poitns in center
grid <- st_make_grid(convex.buff, cellsize = 10e3, what = "centers") %>%
  st_intersection(convex.buff)

## change back to df
bs.grid <- do.call(rbind, st_geometry(grid)) %>%
  as_tibble() %>% setNames(c("X", "Y")) %>%
  # convert df m back to km
  mutate(X = X/1000, Y = Y/1000)
plot(bs.grid)

## replicate df across years and include any other variables we inlcude in our fit of the model
## i.e., depth tmeperature etc
scaled_median <- unique(pcod_lngs_utm$scaled_median)

bs.grid.year <- replicate_df(bs.grid, "sample_year", as.factor(unique(pcod_lngs$sample_year)))
bs.grid.doy <- replicate_df(bs.grid.year, "scale_doy2", scaled_median) # between 235 and 236 is ~0


# maps some of the sites
# ggplot() +
#   geom_sf(data = pcod.sf.shift, aes()) +
#   geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
#   geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
#                breaks = c(-50, -70, -150), color = "black") +
#   #use the 360 coords for longitude (so -150 + 360)
#   coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F)
```

## predict & map predictions 
(both average spatial field but also annual estimates)

```{r}
# predictiosn would be good for fitting the average spatial field 
# prediction here is in link space
predictions <- predict(fit, newdata = bs.grid.doy) %>%
  mutate(Y_m = Y * 1000,
         X_m = X * 1000) # sf requires meters

predictions.sf <- st_as_sf(predictions, coords = c("X_m", "Y_m"), crs = 32603) %>%
  st_transform(crs = 4326) %>%
  st_shift_longitude()

ggplot(predictions, aes(X_m, Y_m, fill = exp(est))) +
  scale_fill_viridis_c() + geom_tile() +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude")

# needs to be exp(est) because you logtransformed your corrected length
lng.by.year <- ggplot() +
  geom_sf(data = predictions.sf, aes(fill = exp(est), color = exp(est)), pch = 22) +
  scale_color_viridis_c() + scale_fill_viridis_c() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  facet_wrap(~as.factor(sample_year)) +
  labs(x = "Longitude", y = "Latitude", color = "Predicted length (mm)",
       fill = "Predicted length (mm)") +
  theme(text = element_text(size = 16))

# average spatial field
lng.avg.spatial <- ggplot() +
  geom_sf(data = predictions.sf, aes(fill = exp(omega_s), color = exp(omega_s)
                                     ), pch = 22) +
  scale_color_viridis_c() + scale_fill_viridis_c() +
  geom_sf(data = pcod_lngs_sf, aes()) +
  geom_sf(data = bs_base, fill = "gray", color = "black", inherit.aes = F) +
  geom_contour(data = bathline.xyz, aes(x = V1 + 360, y = V2, z = V3),
               breaks = c(-50, -70, -150), color = "black") +
  # use the 360 coords for longitude (so -150 + 360)
  coord_sf(xlim = c(180, 202), ylim = c(54, 66), expand = F) +
  labs(x = "Longitude", y = "Latitude", color = "Predicted length (mm)",
       fill = "Predicted length (mm)") +
  theme(text = element_text(size = 16))


# ggsave(filename = "figures/predicted_lng_yearmap_050626.png", plot = lng.by.year,
#        height = 14, width = 14)
# 
# ggsave(filename = "figures/pred_lng_red_avg_spatialmap_072826.png", plot = lng.avg.spatial,
#        height = 14, width = 14)
```


## investigate avg length by year
```{r}
predictions_tmb <- predict(fit,
                           newdata = bs.grid.doy, return_tmb_object = T)

vr <- visreg(fit, xvar = "sample_year", 
             cond=list(scale_doy2 = scaled_median), # this means 238 just in the scaled form
             plot = FALSE, 
             scale = "response")



# partials<-vr$res
# partials$SPECIES<-sp
# fit<-vr$fit
# fit$SPECIES<-sp   




visreg.se.fits<-  vr$fit %>%
  mutate(sample_year = as.character(sample_year),
         sample_year = as.numeric(sample_year)) %>%
  complete(sample_year = 2003:2025) %>%
  mutate(SE = (visregUpr-visregLwr)/ 3.919928)

pred.length <- visreg.se.fits %>%
  ggplot() +
  geom_point(aes(x = sample_year, y = visregFit)) +
  geom_errorbar(aes(x = sample_year, y = visregFit,
                      ymin = visregLwr, ymax = visregUpr), width = 0.2) +
  theme_bw() +
  theme(text = element_text(size = 18)) +
  labs(x = "Year", y = "Predicted length (mm, August 26)", title = "Reduced dataset")

pred.length.se <- visreg.se.fits %>%
  ggplot() +
  geom_point(aes(x = sample_year, y = visregFit)) +
  geom_errorbar(aes(x = sample_year, y = visregFit,
                      ymin = visregFit - SE, ymax = visregFit + SE), width = 0.2) +
  theme_bw() +
  theme(text = element_text(size = 18)) +
  labs(x = "Year", y = "Predicted length (mm, August 31)")

# ggsave(filename = "figures/pred_red_lng_ts_072826.png", plot = pred.length,
#        width = 12, height = 10)

# ggsave(filename = "figures/pred_lng_ts-se_061726.png", plot = pred.length,
#        width = 12, height = 10)
```

What would a temperature model look like? 
```{r}
# this is the annual mean sst for june, july, august with the sd and Z scored values
sst <- read.csv(here("data", "summerSST_BeringSea.csv")) %>%
  filter(region == "EBS", # only want whole EBS since we're modeling it together here
         YearVar > 2002) %>%  # only 2003 to recent 
  mutate(YearVar = as.factor(YearVar)) %>%
  dplyr::select(-X)
```

# Temperature models

We want to understand if there's a relationship with temperature after accounting for spatial variability in lengths (i think).
This is sst that Mariela provided - I believe its surface temperature (aka no depth integrated) so likley represents the average tmep for the surface. 

## fit model
uses same mesh (cause same data)
```{r}
pcod_temp <- pcod_lngs_utm %>%
  left_join(sst, by = c("sample_year" = "YearVar")) %>%
  #clean up old columns
  dplyr::select(-c(resid:region.y))

fit.temp <- sdmTMB(data = pcod_temp,
                   corrected_lng ~ s(scale_doy2) + summerSST_Z,
                    mesh = mesh,
                    time = "sample_year",
                    family = gaussian(link = "log"),
                    spatial = "on",
                    spatiotemporal = "IID",
                    anisotropy = TRUE
                   )

```

## model evaluation
```{r}
## use simulated residuals
# can also look at the simulated residuals
temp.sims <- simulate(fit.temp, nsim = 500, type = "mle-mvn")
temp.resids <- dharma_residuals(temp.sims, fit.temp, return_DHARMa = TRUE)
p_qq3 <- as.grob(~DHARMa::plotQQunif(temp.resids))
p_res3 <- as.grob(~DHARMa::plotResiduals(temp.resids))
  
# combine residual plots
dharma_plots3 <- wrap_plots(p_qq3, p_res3) +
plot_annotation(title = paste("Length red temp DHARMa Diagnostics"))
# ggsave(plot = dharma_plots3, filename = here("figures", "pcod_red_lng_temp_diagnostics.png"))

# spatial residuals
pcod_lngs_utm$scl_temp_resid <- temp.resids$scaledResiduals

# spatial residual plots
spatials3 <- ggplot(pcod_lngs_utm) +
  geom_point(aes(X,Y, col = scl_temp_resid)) +
  labs(title = paste("Pcod red lng temp Spatial residuals")) +
  scale_colour_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  facet_wrap(~sample_year) +
  coord_fixed() +
  theme_minimal()

#ggsave(plot = spatials3, filename = here("figures", "pcod_red_lng_temp_spatial_diagnostics.png"))

```

## prediction grid
temp prediction grid
```{r}
# basically 
## replicate df across years and include any other variables we inlcude in our fit of the model
## i.e., depth tmeperature etc
scaled_median <- unique(pcod_lngs_utm$scaled_median)
mean_tempZ <- mean(pcod_temp$summerSST_Z)
mean_temp <- mean(pcod_temp$summerSST_annualMean) # mean temp is ~7C

bs.grid.doy <- replicate_df(bs.grid.year, "scale_doy2", scaled_median) # between 235 and 236 is ~0
# unsure but i dont think we need to replicate by year because we dont have year in the main effects
bs.grid.temp <- replicate_df(bs.grid.doy, "summerSST_Z", pcod_temp$summerSST_Z)
```


## predict
```{r}
summerSST_Z <- pcod_temp$summerSST_Z
vr.temp <- visreg(fit.temp, xvar = "summerSST_Z", 
             cond=list(scale_doy2 = scaled_median), # this means 238 just in the scaled form
             plot = FALSE, 
             scale = "response")

predict(fit.temp, newdata = data.frame(summerSST_Z = seq(min(pcod_temp$summerSST_Z, na.rm = TRUE),
                                                        max(pcod_temp$summerSST_Z, na.rm = TRUE), 
                                                        length.out = 100),
                                      scale_doy2 = scaled_median),
       )

vr<-visreg(fit.temp, xvar = "summerSST_Z",
           cond=list(scale_doy2=scaled_median),
           plot=FALSE,scale="response")

visreg(fit.temp)
```

